---
title: "HW2"
author: "Moosun Brad Kim"
date: "7/24/2018"
output:
  html_document: default
  pdf_document: default
---

### R fundamentals 03: Elements of programming, R Markdown  (`r format(Sys.Date(), "%b %d, %y")`)
---


---

In this expression below, we can set the default chunk options for the rest of the document
Please note that this chunk itself is defined as `include=FALSE`, means, please evaluate it but no output of any type please!

library(dplyr)
library(knitr)

```{r setDefault, include=FALSE}
# This code chunk defines the default chunks and libary settings
knitr::opts_chunk$set(eval=FALSE, comment="$$$", fig.width=6)
library(dplyr)
library(knitr)
library(rmarkdown)
install.packages("tidyverse")
library(tidyverse)
```


### Data science fundamentals 01: Visualize and explore 
HW2

<span style="color:blue">**Classwork/Homework**</span>: 

**1. Run ggplot(data = HANES). What do you see?**
<span style="color:red">
an empty plot is given since R needs Geom functions or aethetics parameters
</span>
<br>
**2. How many rows are in HANES? How many columns?**
<span style="color:red">
23x1527
</span>
<br>
**3. What does the DX_DBTS variable describe?**
<span style="color:red">
Can map the colors of Diabetes Status to points. Here, diabetes Status is indicated on the legend.
</span>
<br>
**4. Make a scatterplot of HDL vs A1C.**
<span style="color:red">
</span>
<br>

```{r eval=TRUE, comment="Console says:"}
# Load the package RCurl
library(RCurl)
# Import the HANES data set from GitHub; break the string into two for readability
# (Please note this readability aspect very carefully)
URL_text_1 <- "https://raw.githubusercontent.com/kannan-kasthuri/kannan-kasthuri.github.io"
URL_text_2 <- "/master/Datasets/HANES/NYC_HANES_DIAB.csv"
# Paste it to constitute a single URL 
URL <- paste(URL_text_1,URL_text_2, sep="")
HANES <- read.csv(text=getURL(URL))
  # Rename the GENDER factor for identification
  HANES$GENDER <- factor(HANES$GENDER, labels=c("M","F"))
  # Rename the AGEGROUP factor for identification
HANES$AGEGROUP <- factor(HANES$AGEGROUP, labels=c("20-39","40-59","60+"))
  # Rename the HSQ_1 factor for identification
HANES$HSQ_1 <- factor(HANES$HSQ_1, labels=c("Excellent","Very Good","Good", "Fair", "Poor"))
  # Rename the DX_DBTS as a factor
HANES$DX_DBTS <- factor(HANES$DX_DBTS, labels=c("DIAB","DIAB NO_DX","NO DIAB"))
  # Omit all NA from the data frame
HANES <- na.omit(HANES)
  # Observe the structure
str(HANES)
# Load the tidyverse library
library(tidyverse)
  # Make a ggplot
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR)))

ggplot(data = HANES)
str(HANES)

  # Make a ggplot with asthetic color for the variable DX_DBTS
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=DX_DBTS))

##**4. Make a scatterplot of HDL vs A1C.**
ggplot(data = HANES) + 
geom_point(mapping = aes(x = HDL, y = A1C, color="blue"))
```

##

<br>

Classwork/Homework:

What’s gone wrong with this code? Why are the points not blue?
```{r eval=FALSE}
   # Make a ggplot with asthetic shape for the variable DX_DBTS
  ggplot(data = HANES) + 
  geom_point(mapping = aes(x = log(A1C), y = log(UACR), color="blue"))
```
Which variables in HANES are categorical? Which variables are continuous? How can we see this information?
Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
<br>

<span style="color:red">**Work**</span>: 
```{r eval=TRUE, comment="Console says:"}
# Load the package RCurl
library(RCurl)
# Import the HANES data set from GitHub; break the string into two for readability
# (Please note this readability aspect very carefully)
URL_text_1 <- "https://raw.githubusercontent.com/kannan-kasthuri/kannan-kasthuri.github.io"
URL_text_2 <- "/master/Datasets/HANES/NYC_HANES_DIAB.csv"
# Paste it to constitute a single URL 
URL <- paste(URL_text_1,URL_text_2, sep="")
HANES <- read.csv(text=getURL(URL))
  # Rename the GENDER factor for identification
  HANES$GENDER <- factor(HANES$GENDER, labels=c("M","F"))
  # Rename the AGEGROUP factor for identification
HANES$AGEGROUP <- factor(HANES$AGEGROUP, labels=c("20-39","40-59","60+"))
  # Rename the HSQ_1 factor for identification
HANES$HSQ_1 <- factor(HANES$HSQ_1, labels=c("Excellent","Very Good","Good", "Fair", "Poor"))
  # Rename the DX_DBTS as a factor
HANES$DX_DBTS <- factor(HANES$DX_DBTS, labels=c("DIAB","DIAB NO_DX","NO DIAB"))
  # Omit all NA from the data frame
HANES <- na.omit(HANES)
  # Observe the structure
str(HANES)
# Load the tidyverse library
library(tidyverse)
  # Make a ggplot
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR)))

ggplot(data = HANES)
str(HANES)

  # Make a ggplot with asthetic color for the variable DX_DBTS
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=DX_DBTS))

## incorrect ggplot with color inside of aes(). (USE FOR Coloring Cat. Variables)
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color="blue"))
## CORRECT ggplot with color OUTside of aes()
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR)), color="blue")

### 4.Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
#color to continuous variable
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=GLUCOSE))

#ggplot(data = HANES) + 
#geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=HSQ_1))
#
#ggplot(data = HANES) + 
#geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=AGEGROUP))
#
##size to continuous variable
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), size=GLUCOSE))
##shape to continuous variable with color gradient and regtangle shape
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=GLUCOSE), shape=23, fill="blue")

##shape to continuous variable with NO fill=color gradient and regtangle shape. Graident on borders only
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=GLUCOSE), shape=23, fill="blue")

##shape to continuous variable with color gradient and regtangle shape
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=GLUCOSE), shape=15, fill="blue")

#stroke aesthetic. Modify width of border
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=GLUCOSE), shape=15, stroke=5)
##map color aesthetic to a boolien function
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color = A1C < 5))
#
#ggplot(data = HANES) + 
#geom_point(mapping = aes(x = log(A1C), y = log(UACR), size=AGEGROUP))


#ggplot(data = HANES) + 
#geom_point(mapping = aes(x = log(A1C), y = log(UACR), shape=HSQ_1))
#
#ggplot(data = HANES) + 
#geom_point(mapping = aes(x = log(A1C), y = log(UACR), shape=AGEGROUP))

```
<br>

<span style="color:blue">**Classwork/Homework Answers**</span>: 

**1. What’s gone wrong with this code? Why are the points not blue?**
<span style="color:red">
aes() must be closed with no color inside. color= is part of mapping=
</span>
<br>
**2.Which variables in HANES are categorical? Which variables are continuous?**
<span style="color:red">
str(HANES). Or, click on HANES in the environment tab (top-right), and a table representation will open.
</span>
<br>
**3.How can we see this information?**
<span style="color:red">
click on HANES in the environment tab (top-right), and a table representation will open.
</span>
<br>
**4.Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?**
<span style="color:red">
For color, different hues indicate different levels. Shapes does not work for some continuous variables however
</span>
**5. What happens if we map the same variable to multiple aesthetics?**
<span style="color:red">
Can adjust fill, shape, size of geometric points.
</span>
**6. What does the stroke aesthetic do? What shapes does it work with?**
<span style="color:red">
Use the stroke aesthetic to modify the width of the border
</span>
**7. What happens if we map an aesthetic to something other than a variable name, like aes(colour = A1C < 5)?**
<span style="color:red">
Maps color aesthetic to a boolien function. True or False in this case for A1C<5
</span>

<span style="color:red">**Work**</span>: 
##

<br>

Classwork/Homework:

What’s gone wrong with this code? Why are the points not blue?
```{r eval=FALSE}
   # Make a ggplot with asthetic shape for the variable DX_DBTS
  ggplot(data = HANES) + 
  geom_point(mapping = aes(x = log(A1C), y = log(UACR), color="blue"))
```
Which variables in HANES are categorical? Which variables are continuous? How can we see this information?
Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
<br>
Lecture_05_Presentation Slide 43
<span style="color:red">**Work for this HW**</span>: 
```{r eval=TRUE, comment="Console says:"}
# Load the package RCurl
library(RCurl)
# Import the HANES data set from GitHub; break the string into two for readability
# (Please note this readability aspect very carefully)
URL_text_1 <- "https://raw.githubusercontent.com/kannan-kasthuri/kannan-kasthuri.github.io"
URL_text_2 <- "/master/Datasets/HANES/NYC_HANES_DIAB.csv"
# Paste it to constitute a single URL 
URL <- paste(URL_text_1,URL_text_2, sep="")
HANES <- read.csv(text=getURL(URL))
  # Rename the GENDER factor for identification
  HANES$GENDER <- factor(HANES$GENDER, labels=c("M","F"))
  # Rename the AGEGROUP factor for identification
HANES$AGEGROUP <- factor(HANES$AGEGROUP, labels=c("20-39","40-59","60+"))
  # Rename the HSQ_1 factor for identification
HANES$HSQ_1 <- factor(HANES$HSQ_1, labels=c("Excellent","Very Good","Good", "Fair", "Poor"))
  # Rename the DX_DBTS as a factor
HANES$DX_DBTS <- factor(HANES$DX_DBTS, labels=c("DIAB","DIAB NO_DX","NO DIAB"))
  # Omit all NA from the data frame
HANES <- na.omit(HANES)
  # Observe the structure
str(HANES)
# Load the tidyverse library
library(tidyverse)
  # Make a ggplot
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR)))

ggplot(data = HANES)
str(HANES)

  # Make a ggplot with asthetic color for the variable DX_DBTS
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=DX_DBTS))

## incorrect ggplot with color inside of aes(). (USE FOR Coloring Cat. Variables)
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color="blue"))
## CORRECT ggplot with color OUTside of aes()
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR)), color="blue")

### 4.Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
#color to continuous variable
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=GLUCOSE))

#ggplot(data = HANES) + 
#geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=HSQ_1))
#
#ggplot(data = HANES) + 
#geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=AGEGROUP))
#
##size to continuous variable
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), size=GLUCOSE))
##shape to continuous variable with color gradient and regtangle shape
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=GLUCOSE), shape=23, fill="blue")

##shape to continuous variable with NO fill=color gradient and regtangle shape. Graident on borders only
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=GLUCOSE), shape=23, fill="blue")

##shape to continuous variable with color gradient and regtangle shape
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=GLUCOSE), shape=15, fill="blue")

#stroke aesthetic. Modify width of border
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color=GLUCOSE), shape=15, stroke=5)
##map color aesthetic to a boolien function
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR), color = A1C < 5))
#

# Make a ggplot with facet grid - AGEGROUP vs DX_DBTS
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR))) + 
facet_grid(AGEGROUP ~ DX_DBTS)

#ggplot(data = HANES) + 
#geom_point(mapping = aes(x = log(A1C), y = log(UACR), size=AGEGROUP))


#ggplot(data = HANES) + 
#geom_point(mapping = aes(x = log(A1C), y = log(UACR), shape=HSQ_1))
#
#ggplot(data = HANES) + 
#geom_point(mapping = aes(x = log(A1C), y = log(UACR), shape=AGEGROUP))


```


What happens if you facet on a continuous variable?
What plots does the following code make? What does . do?
<br>

<span style="color:blue">**Classwork/Homework Answers**</span>: 

**1. What happens if you facet on a continuous variable?**
<span style="color:red">
Chunk just stops. Cannot facet continuous variables
</span>
<br>
```{r eval=FALSE,results='hold'}
# Make a ggplot with facet grid - CREATININE~ TRIGLYCERIDE
ggplot(data = HANES) + 
geom_point(mapping = aes(x = log(A1C), y = log(UACR))) + 
facet_grid(CREATININE~ TRIGLYCERIDE)
```
<br>
**2.What plots does the following code make? What does . do?**

```{r eval=FALSE}
 ggplot(data = HANES) + 
  geom_point(mapping = aes(x = log(HDL), y = log(CHOLESTEROLTOTAL))) +
  facet_grid(AGEGROUP ~ .)

  ggplot(data = HANES) + 
  geom_point(mapping = aes(x = log(UALBUMIN), y = log(GLUCOSE))) +
  facet_grid(. ~ DX_DBTS)
```
<span style="color:red">
facet_grid(. ~ categorical variable you want with labels on top row)
facet_grid(categorical variable you want with labels on left column~ .)
think as facet_grid(rows~columns). . can be thought of as an empty place holder for the grid.
</span>
<br>
**3. Derive directed insights from the above plots.**
<span style="color:red">

</span>
<br>
**4. Take the first faceted plot in this section:**
```{r eval=TRUE, include=FALSE, comment="Brad says:"}
   # Make a ggplot with facets
  ggplot(data = HANES) + 
  geom_point(mapping = aes(x = log(A1C), y = log(UACR))) + 
  facet_wrap(~ DX_DBTS, nrow = 2)
```
##
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

<span style="color:red">
Sometimes, colors can obsure data points, making it hard to compare. facet_grid lets you visually compare depending on categorical variables. However, colors aesthetics are useful for visualizing levels of continuous variables. For larger datasets with few categorical variables, facet_grid may be an option to compare the combinations of the categories. Color aesthetics is useful if all the categories wish to be seen simultaneously on one graph. 
</span>

**5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol argument?**
<br>
<span style="color:red">
Number of rows and columns. facet_wrap() is generally a better use of screen space than facet_grid() because most displays are roughly rectangular. facet_grid() is most useful when you have two discrete variables, and all combinations of the variables exist in the data.
</span>
<br>
##
**6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?**
<br>
<span style="color:red">
Easier to horizontally read the names of the levels in the columns, and visualize the values continuous variable on rows(vertically)
</span>
<br>
##
---

#### Selected materials and references

1. [An Introduction to R](https://cran.r-project.org/doc/manuals/R-intro.pdf)
2. [knitr package options](https://yihui.name/knitr/options/) and more

---
